---
title: "Homework 3"
author: "Tongtong Zhu"
date: "2022-10-13"
output: github_document
---

```{r setup, include = FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)

library(p8105.datasets)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	message = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


## Problem 1

### Load and read `instacart` data

```{r}
data("instacart")

instacart = 
  instacart %>% 
  as_tibble(instacart)
```

**Description of data**

The `instacart` is a online grocery shopping dataset in 2017. It contains `r nrow(instacart)` observations and `r ncol(instacart)` variables. The key variables include `r colnames(instacart)`.


### Comment on the results

#### Aisle count and identify 

```{r}
instacart %>% 
  group_by(aisle_id, aisle) %>% 
  summarize(n_obs = n()) %>% 
  arrange(desc(n_obs))
  
```

There are `r max(pull(instacart, aisle_id))` aisles and `fresh vegetables`, `fresh fruits` and `packaged vegetables fruits` are the first, second and third aisles with the most items ordered from, respectively. 

#### Plot aisles

```{r}
instacart %>% 
  group_by(aisle) %>% 
  summarize(n_obs = n()) %>% 
  filter(n_obs > 10000) %>% 
  mutate(aisle = forcats::fct_reorder(aisle, n_obs)) %>% 
  ggplot(aes(x = n_obs, y = aisle)) +
  geom_col() +
  labs(
    title = "Number of items ordered per aisle",
    x = "Number of items",
    y = "Aisle name",
    caption = "Note: Limited to aisles with more than 10000 items ordered"
  )

```

#### Table three aisles with three most popular items

```{r}
instacart %>% 
  filter(aisle == "baking ingredients" |
         aisle == "dog food care" |
         aisle == "packaged vegetables fruits") %>% 
  group_by(aisle, product_name) %>% 
  summarize(n_pop = n()) %>% 
  mutate(rank = min_rank(desc(n_pop))) %>% 
  filter(rank < 4) %>% 
  arrange(desc(n_pop)) %>% 
  knitr::kable()
  
```

In `baking ingredients` aisle, the three most popular items are `Light Brown Sugar`, `Pure Baking Soda` and `Cane Sugar`. In `dog food care` aisle, the three most popular items are ` Snack Sticks Chicketn & Rice Recipe Dog Treats` and `Small Dog Biscuits`. In `packaged vegetables fruits` aisle, the three most popular itesm are `Organic Baby Spinach`, `Organic Raspberries` and `Organic Blueberries`.




